---
title: "Tile Data"
author: "Madigan Boborci"
format: html
---


### Load Libraries

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(lubridate)
library(lme4)
library(lmerTest) 
```

### Read in Data Files

```{r}
HOBO1<-read.csv("Data/22263995_08_12.csv")
HOBO2<-read.csv("Data/22263996_08_12.csv")
HOBO3<-read.csv("Data/22263998_08_12.csv")
HOBO4<-read.csv("Data/22263999_08_12.csv")
HOBO5<-read.csv("Data/22264000_08_12.csv")
HOBO6<-read.csv("Data/22264001_08_12.csv")
HOBO7<-read.csv("Data/22264004_08_12.csv")
HOBO8<-read.csv("Data/22264003_08_12.csv")
```

### Add HOBO ID Numbers

```{r}
HOBO1<-mutate(HOBO1, HOBO_ID = "22263995")
HOBO2<-mutate(HOBO2, HOBO_ID = "22263996")
HOBO3<-mutate(HOBO3, HOBO_ID = "22263998")
HOBO4<-mutate(HOBO4, HOBO_ID = "22263999")
HOBO5<-mutate(HOBO5, HOBO_ID = "22264000")
HOBO6<-mutate(HOBO6, HOBO_ID = "22264001")
HOBO7<-mutate(HOBO7, HOBO_ID = "22264004")
HOBO8<-mutate(HOBO8, HOBO_ID = "22264003")
```


#### Use Bind Rows to combine Data

```{r}
HOBO_data_full<- bind_rows(HOBO1, HOBO2, HOBO3, HOBO4, HOBO5, HOBO6, HOBO7, HOBO8)
```


### Clean Data: removing unecessary columns, renaming columns, adding treatments, fixing Date/Time

```{r}
HOBO_clean<- HOBO_data_full%>%
  mutate(Temp_C = Temperature....C) %>%
  rename(Date_Time_HST=Date.Time..HST.)%>%
  mutate(
    Treatment = case_when(
      HOBO_ID %in% c("22264004", "22263998", "22263996", "22264003") ~ "White",
      HOBO_ID %in% c("22263995", "22263999", "22264000","22264001") ~ "Black",
      TRUE                                              ~ "Unknown" 
    ))%>%
      separate(`Date_Time_HST`, into = c("Date", "Time"), sep = " ", remove = FALSE)

write_csv(HOBO_clean, "MFB_HOBO_Clean.csv")
  
```


```{r}
Tides<- read.csv("Data/Raw_Tides_Minute.csv")
```

```{r}
Tides <- Tides %>%
  mutate(
    Date_Time_HST = ymd_hm(paste(Date, Time), tz = "HST")
  )
```

```{r}
HOBOTide<-HOBO_clean %>%
  mutate(
    Date_Time_HST = mdy_hms(Date_Time_HST, tz = "HST")
  )
```

```{r}
HOBO_Tides<- HOBOTide %>%
  left_join(Tides%>%
              select(-Date, -Time,-Zone),
             by="Date_Time_HST")
```


### Plot Data

```{r}
ggplot(data=HOBO_clean, aes(x=Date, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,28)
```


#### Try to isolate Daily Data to see patterns

```{r}
HOBO_OneDay<-HOBO_clean %>%
  filter(Date == "07/02/2026")
```


### Plot that Data

```{r}
ggplot(data=HOBO_OneDay, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

## Try to isolate a smaller date window to observe

#### Going Monthly

```{r}
HOBO_June<- HOBO_clean %>% 
  filter(Date >= "06/08/2026" & Date <= "06/30/2026")
```

```{r}
ggplot(data=HOBO_June, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

```{r}
HOBO_July<- HOBO_clean %>% 
  filter(Date >= "07/01/2026" & Date <= "07/31/2026")
```

```{r}
ggplot(data=HOBO_July, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

```{r}
HOBO_August<- HOBO_clean %>% 
  filter(Date >= "08/01/2026" & Date <= "08/12/2026")
```

```{r}
ggplot(data=HOBO_August, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

# T tests and comparisons 

```{r}
t_result <- t.test(Temp_C ~ Treatment, data = HOBO_clean)

t_result
```

```{r}
avg_tile<- HOBO_clean %>%
  group_by(Treatment)%>%
  summarize(mean(Temp_C))
avg_tile
```



### Starting Fresh
```{r}
daily_max<-HOBO_clean%>%
  group_by(Date, Treatment, HOBO_ID) %>%
  summarise(
    daily_max = max(Temp_C, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  mutate(
    daily_max = ifelse(is.infinite(daily_max), NA, daily_max)
  )
```
```{r}
t_daily_max <- t.test(daily_max ~ Treatment, data = daily_max)
```


```{r}
anova(lm(daily_max~Treatment, data=daily_max))
```
## This is me trying multiple ways hoping to get a better result 

```{r}
daily_max%>%
  group_by(Treatment)%>%
  summarise(
    avg_daily_max = mean(daily_max, na.rm = TRUE), 
    .groups = "drop"
  )
```
# Good Graph
```{r}
ggplot(data=HOBO_clean, aes(x=Date_Time_HST, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,40)
ggsave("June-Aug.png")
```
```{r}
HOBO_clean <- HOBO_clean %>%
  mutate(Date_Time_HST = mdy_hms(Date_Time_HST))
```



# Do a max difference comparison 

```{r}
daily_max_df <- HOBO_clean %>%
  filter(!is.na(Temp_C)) %>%
  # Use as_date() directly from lubridate
  mutate(Date = as_date(Date_Time_HST)) %>% 
  group_by(Date, HOBO_ID, Treatment) %>%
  summarise(
    max_temp = max(Temp_C, na.rm = TRUE),
    .groups = "drop"
  )
```


```{r}
max_model <- lmer(max_temp ~ Treatment + (1 | HOBO_ID), data = daily_max_df)
summary(max_model)
```

```{r}
max_t_test<-daily_max_df %>%
  group_by(HOBO_ID, Treatment) %>%
  summarise(mean_daily_max = mean(max_temp, na.rm = TRUE), .groups = "drop")

t.test(mean_daily_max ~ Treatment, data = max_t_test)
```

```{r}
treatment_daily_summary <- daily_max_df %>%
  group_by(Date, Treatment) %>%
  summarise(avg_max_temp = mean(max_temp, na.rm = TRUE), .groups = "drop") %>%
  # Pivot treatments into columns to compute the difference
  pivot_wider(names_from = Treatment, values_from = avg_max_temp) %>%
  mutate(
    # Absolute difference between black and white treatment averages
    temp_diff = abs(Black - White),
    # Signed difference (positive means Black was hotter, negative means White was hotter)
    black_minus_white = Black - White
  )
```

```{r}
avg_offset <- mean(treatment_daily_summary$black_minus_white, na.rm = TRUE)
avg_offset
```


```{r}
# Step B: Find the day with the LARGEST temperature gap
largest_gap <- treatment_daily_summary %>%
  filter(!is.na(temp_diff)) %>%
  slice_max(order_by = temp_diff, n = 1)

# View the result
print("Day with the largest temperature difference between treatments:")
print(largest_gap)
```

```{r}
ggplot(treatment_daily_summary, aes(x = Date, y = black_minus_white)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = largest_gap, aes(x = Date, y = black_minus_white), color = "red", size = 4) +
  annotate(
    "text", 
    x = largest_gap$Date, 
    y = largest_gap$black_minus_white, 
    label = paste0(" Max Gap: ", round(largest_gap$temp_diff, 2), "°C"),
    hjust = -0.1, vjust = -0.5, fontface = "bold"
  ) +
  labs(
    title = "Daily Temperature Difference (Black Treatment - White Treatment)",
    subtitle = "Positive values indicate Black treatment was hotter",
    x = "Date",
    y = "Temperature Difference (°C)"
  ) +
  theme_minimal()

ggsave("Temp_Diff_Graph.png")
```

# Tide Correcting 

add in blocks for each hobo and then add what the tide height is for each block and then do multiple if else statements to account for each block

Then group by treatment to get the average exposed/submerged

if block height >tide height then exposed=True
if block height <tide height then exposed= false

for mica- then group by treatment and average for exposed=false


```{r}
HOBO_Tides<-HOBO_Tides%>%
  mutate(
    Block = case_when(
      HOBO_ID %in% c("22264004", "22263995") ~ "2",
      HOBO_ID %in% c("22263998", "22264000") ~ "4",
      HOBO_ID %in% c("22263999", "22263996") ~ "6",
      HOBO_ID %in% c("22264001", "22264003") ~ "8",
      TRUE                                              ~ "Unknown" 
    ))
```


if tide_height > 0.26 for Block 2 then submerged=TRUE
if tide_height > 0.05 for Block 4 then submerged=TRUE
if tide_height > 0.003 for Block 6 then submerged=TRUE
if tide_height > 0.065 for Block 8 then submerged=TRUE

```{r}
HOBO_Tides<- HOBO_Tides%>%
  mutate(result = case_when(
    Block == "2" & Tide_Height > 0.26  ~ "Submerged",
    Block == "4" & Tide_Height > 0.05 ~ "Submerged",
    Block == "6" & Tide_Height > 0.003 ~ "Submerged",
    Block == "8" & Tide_Height > 0.065 ~ "Submerged",
    TRUE ~ "Exposed" 
  ))
```

```{r}
avg<- HOBO_Tides%>%
  group_by(Treatment, result)%>%
  summarize(mean(Temp_C),
            min(Temp_C),
            max(Temp_C))
```

```{r}
HOBO_Tides%>%
  filter(result== "Exposed" & Treatment== "Black")

%>%
  summarize(mean(Temp_C))
```


For showing Black Tiles work: take exposed days find the max temps and then average those across time












# random 
```{r}
HOBO_clean <- HOBO_clean %>%
  mutate(Date_Time_HST = mdy_hms(Date_Time_HST))
```

```{r}
test <- HOBO_clean %>%
  filter(
    Date_Time_HST >= mdy_hms("07/01/2026 01:00:00"),
    Date_Time_HST <= mdy_hms("07/04/2026 23:00:00")
  )
```

```{r}
ggplot(data = test, aes(x = Date_Time_HST, y = Temp_C, color = factor(HOBO_ID))) +
  geom_point() +
  geom_line()
```


# Not sure if any of this was right for ADM- relook at it


group by day, get max temp and average through the month ~Treatment 



adm_ttest <- t.test(daily_max ~ Treatment, data = adm)





adm_July<-HOBO_July%>%
  group_by(Date, Treatment) %>%
  summarise(
    daily_max = max(Temp_C, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  group_by(Treatment) %>%
  summarise(
    avg_daily_max = mean(daily_max, na.rm = TRUE), 
    .groups = "drop"
  )
  



adm_June_for_ttest <- HOBO_June %>%
  group_by(Date, Treatment) %>%
  summarise(
    daily_max = max(Temp_C, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  mutate(
    daily_max = ifelse(is.infinite(daily_max), NA, daily_max)
  )


t_test_June <- t.test(daily_max ~ Treatment, data = adm_June_for_ttest)


adm_June_for_ttest %>%
  filter(!is.na(daily_max)) %>%
  group_by(Treatment) %>%
  summarise(sample_size = n())



#### Tide Notes for Winter sites- not concerned about tides not heating tiles enough

By end of December, tides (Diamond Head) are in mid afternoon (1-3)
Before that, there's a higher low in the am about 8 hrs before the low low, usually around the time the summer tides have been in the am, but only at about 0.2-0 as lowest values. Same ish with Sandy's and Ewa with afternoon lows right at 0 by mid-end of Dec. 

Probably want to do tidal heights at the 0.1 end of the tile block spectrum that we have at TB


