---
title: "Tile Data"
author: "Madigan Boborci"
format: html
---


### Load Libraries

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(lubridate)
library(lme4)
library(lmerTest) 
```

### Read in Data Files

```{r}
HOBO1<-read.csv("Data/22263995_20260728.csv")
HOBO2<-read.csv("Data/22263996 2026-07-12 06_15_15 HST.csv")
HOBO3<-read.csv("Data/22263998_20260728 .csv")
HOBO4<-read.csv("Data/22263999_20260728 .csv")
HOBO5<-read.csv("Data/22264000_20260728 .csv")
HOBO6<-read.csv("Data/22264001_20260728.csv")
HOBO7<-read.csv("Data/22264004_20260728.csv")
```

### Add HOBO ID Numbers

```{r}
HOBO1<-mutate(HOBO1, HOBO_ID = "22263995")
HOBO2<-mutate(HOBO2, HOBO_ID = "22263996")
HOBO3<-mutate(HOBO3, HOBO_ID = "22263998")
HOBO4<-mutate(HOBO4, HOBO_ID = "22263999")
HOBO5<-mutate(HOBO5, HOBO_ID = "22264000")
HOBO6<-mutate(HOBO6, HOBO_ID = "22264001")
HOBO7<-mutate(HOBO7, HOBO_ID = "22264004")
```


#### Use Bind Rows to combine Data

```{r}
HOBO_data_full<- bind_rows(HOBO1, HOBO2, HOBO3, HOBO4, HOBO5, HOBO6, HOBO7)
```


### Clean Data: removing unecessary columns, renaming columns, adding treatments, fixing Date/Time

```{r}
HOBO_clean<- HOBO_data_full%>%
  select(-`Host.Connected`, -`End.of.File`, -`Button.Down`,-`Button.Up`)%>%
  mutate(Temp_C = coalesce(Temperature.....C., Temperature....C)) %>%
  select(-Temperature.....C., -Temperature....C)%>%
  rename(Date_Time_HST=Date.Time..HST.)%>%
  mutate(
    Treatment = case_when(
      HOBO_ID %in% c("22264004", "22263998", "22263996") ~ "White",
      HOBO_ID %in% c("22263995", "22263999", "22264000","22264001") ~ "Black",
      TRUE                                              ~ "Unknown" 
    ))%>%
      separate(`Date_Time_HST`, into = c("Date", "Time"), sep = " ", remove = FALSE)

write_csv(HOBO_clean, "MFB_HOBO_Clean.csv")
  
```



### Plot Data

```{r}
ggplot(data=HOBO_clean, aes(x=Date, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,28)
```


#### Try to isolate Daily Data to see patterns

```{r}
HOBO_OneDay<-HOBO_clean %>%
  filter(Date == "07/02/2026")
```


### Plot that Data

```{r}
ggplot(data=HOBO_OneDay, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

## Try to isolate a smaller date window to observe

#### Going Monthly

```{r}
HOBO_July<- HOBO_clean %>% 
  filter(Date >= "07/01/2026" & Date <= "07/28/2026")
```

```{r}
ggplot(data=HOBO_July, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

```{r}
HOBO_June<- HOBO_clean %>% 
  filter(Date >= "06/08/2026" & Date <= "06/30/2026")
```

```{r}
ggplot(data=HOBO_June, aes(x=Time, y=Temp_C, color=Treatment))+
                            geom_line()+
  ylim(22,35)
```

# Significant Results!

```{r}
t_result <- t.test(Temp_C ~ Treatment, data = HOBO_clean)

t_result
```

group by day, get max temp and average through the month ~Treatment 



adm_ttest <- t.test(daily_max ~ Treatment, data = adm)




```{r}
adm_July<-HOBO_July%>%
  group_by(Date, Treatment) %>%
  summarise(
    daily_max = max(Temp_C, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  group_by(Treatment) %>%
  summarise(
    avg_daily_max = mean(daily_max, na.rm = TRUE), 
    .groups = "drop"
  )
  
```


```{r}
adm_June_for_ttest <- HOBO_June %>%
  group_by(Date, Treatment) %>%
  summarise(
    daily_max = max(Temp_C, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  mutate(
    daily_max = ifelse(is.infinite(daily_max), NA, daily_max)
  )

```
```{r}
t_test_June <- t.test(daily_max ~ Treatment, data = adm_June_for_ttest)
```

```{r}
adm_June_for_ttest %>%
  filter(!is.na(daily_max)) %>%
  group_by(Treatment) %>%
  summarise(sample_size = n())

```



### Starting Fresh
```{r}
daily_max<-HOBO_clean%>%
  group_by(Date, Treatment, HOBO_ID) %>%
  summarise(
    daily_max = max(Temp_C, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  mutate(
    daily_max = ifelse(is.infinite(daily_max), NA, daily_max)
  )
```
```{r}
t_daily_max <- t.test(daily_max ~ Treatment, data = daily_max)
```


```{r}
anova(lm(daily_max~Treatment, data=daily_max))
```

```{r}
daily_max%>%
  group_by(Treatment)%>%
  summarise(
    avg_daily_max = mean(daily_max, na.rm = TRUE), 
    .groups = "drop"
  )
```
```{r}
ggplot(data=HOBO_clean, aes(x=Date_Time_HST, y=Temp_C, color=Treatment))+
                            geom_line()
```
```{r}
HOBO_clean <- HOBO_clean %>%
  mutate(Date_Time_HST = mdy_hms(Date_Time_HST))
```

```{r}
test <- HOBO_clean %>%
  filter(
    Date_Time_HST >= mdy_hms("07/01/2026 01:00:00"),
    Date_Time_HST <= mdy_hms("07/04/2026 23:00:00")
  )
```

```{r}
ggplot(data = test, aes(x = Date_Time_HST, y = Temp_C, color = factor(HOBO_ID))) +
  geom_point() +
  geom_line()
```

### Do a max difference comparison 














```{r}
daily_max_df <- HOBO_clean %>%
  filter(!is.na(Temp_C)) %>%
  mutate(Date = as.Date(Date_Time_HST)) %>% 
  group_by(Date, HOBO_ID, Treatment) %>%
  summarise(
    max_temp = max(Temp_C, na.rm = TRUE),
    .groups = "drop"
  )
```



```{r}
max_model <- lmer(max_temp ~ Treatment + (1 | HOBO_ID), data = daily_max_df)
summary(max_model)
```

```{r}
max_t_test<-daily_max_df %>%
  group_by(HOBO_ID, Treatment) %>%
  summarise(mean_daily_max = mean(max_temp, na.rm = TRUE), .groups = "drop")

t.test(mean_daily_max ~ Treatment, data = max_t_test)
```

```{r}
treatment_daily_summary <- daily_max_df %>%
  group_by(Date, Treatment) %>%
  summarise(avg_max_temp = mean(max_temp, na.rm = TRUE), .groups = "drop") %>%
  # Pivot treatments into columns to compute the difference
  pivot_wider(names_from = Treatment, values_from = avg_max_temp) %>%
  mutate(
    # Absolute difference between black and white treatment averages
    temp_diff = abs(Black - White),
    # Signed difference (positive means Black was hotter, negative means White was hotter)
    black_minus_white = Black - White
  )
```

```{r}
avg_offset <- mean(treatment_daily_summary$black_minus_white, na.rm = TRUE)
avg_offset
```


```{r}
# Step B: Find the day with the LARGEST temperature gap
largest_gap <- treatment_daily_summary %>%
  filter(!is.na(temp_diff)) %>%
  slice_max(order_by = temp_diff, n = 1)

# View the result
print("Day with the largest temperature difference between treatments:")
print(largest_gap)
```

```{r}
ggplot(treatment_daily_summary, aes(x = Date, y = black_minus_white)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = largest_gap, aes(x = Date, y = black_minus_white), color = "red", size = 4) +
  annotate(
    "text", 
    x = largest_gap$Date, 
    y = largest_gap$black_minus_white, 
    label = paste0(" Max Gap: ", round(largest_gap$temp_diff, 2), "°C"),
    hjust = -0.1, vjust = -0.5, fontface = "bold"
  ) +
  labs(
    title = "Daily Temperature Difference (Black Treatment - White Treatment)",
    subtitle = "Positive values indicate Black treatment was hotter",
    x = "Date",
    y = "Temperature Difference (°C)"
  ) +
  theme_minimal()
```



#### Tide Notes for Winter sites- not concerned about tides not heating tiles enough

By end of December, tides (Diamond Head) are in mid afternoon (1-3)
Before that, there's a higher low in the am about 8 hrs before the low low, usually around the time the summer tides have been in the am, but only at about 0.2-0 as lowest values. Same ish with Sandy's and Ewa with afternoon lows right at 0 by mid-end of Dec. 

Probably want to do tidal heights at the 0.1 end of the tile block spectrum that we have at TB


